Viscous Transport in Eroding Porous Media

### 2019-09-02 docs/transport/erosion_transport.tex
modify something  in transport paper
worked on low porosity 20 body example

In Figure~\ref{fig:Eroding20Transport}(a), we plot the tortuosity as a
function of the porosity from the initial porosity of $0.38$ until all
the bodies have eroded.  The tortuosity is computed with both the length
of the streamlines~\eqref{eqn:tortuosity1} (red marks) and using the
spatial average of the velocity on an Eulerian
grid~\eqref{eqn:tortuosity2} (blue marks).  The red square corresponds
to the porosity of the geometry in Figure~\ref{fig:Eroding20tort}(c).
The two tortuosity formulas give similar results, and any discrepancy
can be accounted for by slow regions of recirculation and quadrature
error induced by the integrals in the tortuosity definitions.  As the
bodies erode, this creates wide channels where streamlines undergo only
minor vertical deflections, and this explains why the tortuosity
eventually decreases with porosity.  We computed lines of best fit using
the porosity-tortuosity models~\eqref{eqn:tortuosityModels}, and the
power law model gave the smallest error.  The  black dashed line in
Figure~\ref{fig:Eroding20Transport}(a) is the line of best fit
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.2064}
\end{align}
with a root-mean-square error of $5.9 \times 10^{-3}$.  The initial
increase in tortuosity occurs because in the absence of erosion (left
plot in Figure~\ref{fig:Eroding20vort}), many of the streamlines, such
as those initialized between bodies 15 \& 18, only perform minor
deflections to pass through the narrow regions, albeit, very slowly.
However, as erosion starts to open the channels, the streamlines deflect
into the fast regions, such as the region above body 11, and this
increases the amount of vertical deflection, and therefore the
tortuosity.

@@

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/tort_eulerian}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/20b_second_moment_long_ref}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20Transport} (a): The tortuosity inside of an
eroding geometry initialized with 20 bodies.  The tortuosity is
calculated using the Eulerian method~\eqref{eqn:tortuosity2} (blue dots)
and Lagrangian method~\eqref{eqn:tortuosity1} (red stars).  The red
square correspond to the geometry in Figure~\ref{fig:Eroding20tort}(c).
The dash line is the line of best fit $\widehat{T}(\phi)=\phi^{-p}$ with
$p=0.2064$, and the corresponding root-mean-square error is $5.9 \times
10^{-3}$. (b): The temporal evolution of $\sigma_\lambda$ as a function
of time at seven different porosities.  The dashed, which has slope one,
corresponds to $\sigma_\lambda \sim t$ if no bodies were present.  At
early times, the particles undergo a ballistic motion, but once they
have traversed a few grains, the dispersion becomes super-dispersive
with $\sigma_\lambda \sim t^\alpha$, $\alpha \in (1/2,1)$.  The
dashed-dotted lines are lines of best fit with slopes $\alpha = 1.06$
($\phi=95.1\%$), $\alpha = 1.07$ ($\phi=85.09\%$), $\alpha = 1.06$
($\phi=75.15\%$), $\alpha = 0.97$ ($\phi=65.09\%$), $\alpha = 0.78$
($\phi=55.10\%$), $\alpha = 0.75$ ($\phi=45.08\%$), and $\alpha = 0.56$
($\phi=37.68\%$).  Values greater than 1 are a result of using a
least-squares fit for the tails of the particle spreading.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Low Porosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider a second example with 20 eroding bodies, but with a smaller
initial porosity.  In Figure~\ref{fig:ErodingLow20vort}, we plot the
eroding geometry and porosity at four evenly spaced instances in time.
Again, the color is the vorticity in the bulk whose magnitude is
equivalent to the erosion rate.   Initially, the smallest distance
between pairs of bodies is $3.29 \times 10^{-4}$, and the smallest
distance between the bodies and solid wall is $4.50 \times 10^{-3}$.  At
these distances, a resolution of approximately $N_\iin = 27,000$ and
$N_\out = 18,000$ discretization points is required to satisfy the $5h$
rule of thumb that is required for the trapezoid rule to achieve machine
epsilon accuracy.  By applying the Barycentric quadrature rule, we
stably simulate erosion at the much coarser resolution of $N_\iin = 256$
and $N_\out = 1024$ discretization points.

\begin{figure}[H]
\begin{center}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort1}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort150}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort300}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort450}
\end{center}
\caption{\label{fig:ErodingLow20vort} Erosion of 20 nearly touching
bodies with the fixed pressure drop condition. The color is the
vorticity of the fluid. The 4 snapshots are evenly spaced in time. We
use $N_\iin = 256$ discretization points on the body, a time-step of
$\Dt=10^{-4}$, and smoothing parameters of $\eps= 15/256$ and
$\sigma=10/256 $. }
\end{figure}

We compute the tortuosity using the Eulerian
method~\eqref{eqn:tortuosity2}. In Figure~\ref{fig:ErodingLow20tort}, we
plot the tortuosity with respect to the porosity (blue) and a line of
best fit (black) using the power law $\widehat{T}(\phi) =
\phi^{-p}$~\cite{mat-kha-koz2008}.  This model outperformed the other
three models described in equation~\eqref{eqn:tortuosityModels}.  The
root-mean-squared error of the power law is $1.13 \times 10^{-2}$.

Unlike the previous 20 body simulation, the tortuosity decreases
monotonically with the porosity. \todo[inline]{BQ: Why is this the case?
What is special about the last example?}

\begin{figure}[H]
\center
\includegraphics*[width =0.55\linewidth]{./figs/tort_eulerian20b}
\caption{\label{fig:ErodingLow20tort} The tortuosity inside of an
eroding geometry initialized with 20 bodies.  The tortuosity is
calculated using the Eulerian method~\eqref{eqn:tortuosity2} (blue
dots).  The dash line is the fitting line $\widehat{T}(\phi)=\phi^{-p}$
as $p=0.1669$ and the root mean square deviation is $1.13 \times
10^{-2}$.}
\end{figure}

\begin{figure}[H]
\center
\includegraphics*[width =0.55\linewidth]{./figs/20b_dense_second_moment_ref}
\caption{\label{fig:Eroding20anomalousLow} The analysis of anomalous
dispersion in 20 nearly touching bodies with respect to initial porosity
(as p=30.67\%) and another seven porosities during the eroding. The dashed line
correspondes to a porosity of 100\% and the dashed-dotted lines are lines
of best fit with slope 0.9194 (as p=95\%), 0.8666 (as p=85.17\%), 
0.8682 (as p=75.10\%), 0.9087 (as p=65.02\%), 
0.9054 (as p=55.08\%), 0.7234 (as p=45.03\%), 
0.6895 (as p=35.05\%), and 0.9225 (as p=30.67\%).}
\end{figure}

@@

>>>>

In Figure~\ref{fig:Eroding20Transport}(a), we plot the tortuosity as a
function of the porosity from the initial porosity of $0.38$ until all
the bodies have eroded.  The tortuosity is computed with both the length
of the streamlines~\eqref{eqn:tortuosity1} (red marks) and using the
spatial average of the velocity on an Eulerian
grid~\eqref{eqn:tortuosity2} (blue marks).  The red square corresponds
to the porosity of the geometry in Figure~\ref{fig:Eroding20tort}(c).
The two tortuosity formulas give similar results, and any discrepancy
can be accounted for by slow regions of recirculation and quadrature
error induced by the integrals in the tortuosity definitions.  As the
bodies erode, this creates wide channels where streamlines undergo only
minor vertical deflections, and this explains why the tortuosity
eventually decreases with porosity.  We computed lines of best fit using
the porosity-tortuosity models~\eqref{eqn:tortuosityModels}, and the
power law model gave the smallest error.  The  black dashed line in
Figure~\ref{fig:Eroding20Transport}(a) is the line of best fit
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.2064}
\end{align}
with a root-mean-square error of $5.90 \times 10^{-3}$.  The initial
increase in tortuosity occurs because in the absence of erosion (left
plot in Figure~\ref{fig:Eroding20vort}), many of the streamlines, such
as those initialized between bodies 15 \& 18, only perform minor
deflections to pass through the narrow regions, albeit, very slowly.
However, as erosion starts to open the channels, the streamlines deflect
into the fast regions, such as the region above body 11, and this
increases the amount of vertical deflection, and therefore the
tortuosity.

@@

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/tort_eulerian}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/20b_second_moment_long_ref}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20Transport} (a): The tortuosity of an
eroding geometry initialized with 20 bodies.  The tortuosity is
calculated using the Eulerian method~\eqref{eqn:tortuosity2} (blue dots)
and Lagrangian method~\eqref{eqn:tortuosity1} (red stars).  The red
square correspond to the geometry in Figure~\ref{fig:Eroding20tort}(c).
The dash line is the line of best fit $\widehat{T}(\phi)=\phi^{-p}$ with
$p=0.2064$, and the corresponding root-mean-square error is $5.90 \times
10^{-3}$. (b): The temporal evolution of $\sigma_\lambda$ as a function
of time at seven different porosities.  The dashed line, which has slope
one, corresponds to the ballistic dispersion $\sigma_\lambda \sim t$ if
no bodies were present. At early times, the particles undergo a
ballistic motion, but once they have traversed a few grains, the
dispersion becomes super-dispersive with $\sigma_\lambda \sim t^\alpha$,
$\alpha \in (1/2,1)$.  The dashed-dotted lines are lines of best fit
with slopes $\alpha = 1.06$ ($\phi=95.10\%$), $\alpha = 1.07$
($\phi=85.09\%$), $\alpha = 1.06$ ($\phi=75.15\%$), $\alpha = 0.97$
($\phi=65.09\%$), $\alpha = 0.78$ ($\phi=55.10\%$), $\alpha = 0.75$
($\phi=45.08\%$), and $\alpha = 0.56$ ($\phi=37.68\%$).  Values greater
than 1 are a result of using a least-squares fit for the tails of the
particle spreading.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Low Porosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider a second example with 20 eroding bodies, but with a smaller
initial porosity.  In Figure~\ref{fig:ErodingLow20vort}, we plot the
eroding geometry and porosity at four evenly spaced instances in time.
Again, the color is the vorticity in the bulk whose magnitude is
equivalent to the erosion rate.   Initially, the smallest distance
between pairs of bodies is $3.29 \times 10^{-4}$, and the smallest
distance between the bodies and solid wall is $4.50 \times 10^{-3}$.  At
these distances, a resolution of approximately $N_\iin = 27,000$ and
$N_\out = 18,000$ discretization points is required to satisfy the $5h$
rule of thumb for the trapezoid rule to achieve machine epsilon
accuracy.  By applying the Barycentric quadrature rule, we stably
simulate erosion at the much coarser resolution of $N_\iin = 256$ and
$N_\out = 1024$ discretization points.

\begin{figure}[H]
\begin{center}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort1}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort150}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort300}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort450}
\end{center}
\caption{\label{fig:ErodingLow20vort} Erosion of 20 nearly touching
bodies with the fixed pressure drop condition. The color is the
vorticity of the fluid. The 4 snapshots are evenly spaced in time. We
use $N_\iin = 256$ discretization points on the bodies.  In
addition to the channels that develop between the bodies and the solid
walls, Erosion leads to two main channels through the geometry---one in
the top half and one near the middle.}
\end{figure}

We compute the tortuosity using the Eulerian
method~\eqref{eqn:tortuosity2}. In
Figure~\ref{fig:ErodingLow20Transport}(a), we plot the tortuosity with
respect to the porosity (blue) and the line of best fit (black) using
the power law 
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.1669}.
\end{align}
This model outperformed the other three models in
equation~\eqref{eqn:tortuosityModels} that were proposed by Matyka et
al.~\cite{mat-kha-koz2008}.  The root-mean-squared error of the power
law is $1.13 \times 10^{-2}$.  As the bodies erode, the number of
streamlines that take a direct path through the geometry increases,
and this decreases the tortuosity.  However, there is also an increase
in the number of streamlines that increase their length since they
deflect their trajectory to move from a low porosity region (high
pressure) to a high porosity region (low pressure), and this decreases
the tortuosity. For this example, we see that the net effect is a
decrease in the tortuosity for all time.

In Figure~\ref{fig:ErodingLow20Transport}(b), we plot the temporal
evolution of the particle spreading $\sigma_\lambda$.  As in the last
example, we analyze the spreading at several different
porosities and we use the reinsertion algorithm described in Section~\ref{sec:dispersion}
to form sufficiently long trajectories.  For all the reported
porosities, the particle dispersion exhibits two power low regimes, with
the lowest porosity geometry, $\phi=37.68\%$ having the clearest
transition.  Again, the dispersion transforms from a ballistic to
super-diffusive, and the anomalous disperion rate increases with the
porosity.  For this example, the dispersion is nearly Fickean at the
smallest porosity.
\todo[inline]{BQ: This paragraph has some false statements.  Need to
look closer at the data.  Might have accidently been looking at the
other 20 body example ie. Fickean dispersion}

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/tort_eulerian20b}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height=0.8\linewidth]{./figs/20b_dense_second_moment_ref}
\caption{}
\end{subfigure}
\caption{\label{fig:ErodingLow20Transport} (a): The tortuosity of an
eroding geometry initialized with 20 bodies.  When compared to the last
example, the bodies are initially much closer together, and the porosity
is smaller.  The tortuosity is
calculated using the Eulerian method~\eqref{eqn:tortuosity2} (blue
dots).  The dash line is the fitting line $\widehat{T}(\phi)=\phi^{-p}$
with $p=0.1669$, and the root mean square deviation is $1.13 \times
10^{-2}$.  (b): The temporal evolution of $\sigma_\lambda$ as a function
of time at eight different porosities.  The dashed line, which has slope
one, corresponds to the ballistic dispersion $\sigma_\lambda \sim t$ if
no bodies were present.  At early times, the particles undergo a
ballistic motion, but once they have traveresd a few grains, the
dispersion becomes super-dispersive with $\sigma_\lambda \sim
t^{\alpha}$, $\alpha \in (1/2,1)$.  The dashed-dotted lines are lines of
best fit with slope $0.92$ ($\phi=95.00\%$), $0.87$ ($\phi=85.17\%$),
$0.87$ ($\phi=75.10\%$), $0.91$ ($\phi=65.02\%$), $0.91$
($\phi=55.08\%$), $0.72$ ($\phi=45.03\%$), $0.69$ ($\phi=35.05\%$), and
$0.92$ ($\phi=30.67\%$).}
\end{figure}

@@

%\begin{figure}[H]
%\center
%\includegraphics*[width =0.55\linewidth]{./figs/tort_eulerian20b}
%\caption{\label{fig:ErodingLow20tort} The tortuosity inside of an
%eroding geometry initialized with 20 bodies.  The tortuosity is
%calculated using the Eulerian method~\eqref{eqn:tortuosity2} (blue
%dots).  The dash line is the fitting line $\widehat{T}(\phi)=\phi^{-p}$
%as $p=0.1669$ and the root mean square deviation is $1.13 \times
%10^{-2}$.}
%\end{figure}
%
%\begin{figure}[H]
%\center
%\includegraphics*[width =0.55\linewidth]{./figs/20b_dense_second_moment_ref}
%\caption{\label{fig:Eroding20anomalousLow} The analysis of anomalous
%dispersion in 20 nearly touching bodies with respect to initial porosity
%(as p=30.67\%) and another seven porosities during the eroding. The
%dashed line correspondes to a porosity of 100\% and the dashed-dotted
%lines are lines of best fit with slope 0.9194 (as p=95\%), 0.8666 (as
%p=85.17\%), 0.8682 (as p=75.10\%), 0.9087 (as p=65.02\%), 0.9054 (as
%p=55.08\%), 0.7234 (as p=45.03\%), 0.6895 (as p=35.05\%), and 0.9225 (as
%p=30.67\%).}
%\end{figure}
